---
title: "Midterm October 15"
author: "Anika Bennett"
date: "2025-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# JOUR405: Statistics for Journalists
## Midterm Exam - Fall 2025

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences.

# Load tidyverse
```{r}
library(tidyverse)
```

## Part 1: Restaurant Health Inspections (25 points)

The first dataset contains restaurant health inspection scores from Montgomery County. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv`

The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible and the compliance score and grade.

## Task 1: Read the data**
```{r}
Health <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```


## Task 2: Calculate the mean and standard deviation maximum and minimum of compliance scores. Put the results in a new dataframe called "compliance_stats." 
```{r}
compliance_stats <- Health |> summarize(mean = mean(compliance_score), sd = sd(compliance_score), max(compliance_score), min(compliance_score))

```

## Task 3: Create a histogram of the compliance scores with a vertical line showing the mean. Include a headline. Include a headline and label the x and y axes.
```{r}
Health |> 
  ggplot(aes(x= compliance_score)) +
  geom_histogram(bins = 10, fill = "blue", color = "black") +
  geom_vline(aes(xintercept = mean(compliance_score)), color = "red", linetype = "dashed") +
  labs(title = "Montgomery County Health Inspection Scores", x = "Compliance Score", y = "Establishment ID")
```

## Task 4: Write 2-3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What would be newsworthy about this distribution? What's the story here? 

**The mean and the standard deviation deviation tell us that the average establishment does align with the health codes, with majority getting A ratings. The story follows the establishments that do not align with health codes and even got those F ratings. Informing the public of the not so safe establishments because of health violations is very important.**


## Part 2: High School Athletics (30 points)

The second dataset shows participation numbers in high school sports across Maryland counties in 2024, broken down by sport and gender. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv`

# Task 5: Load the high school participation data
```{r}
Participation <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv")
```


#Task 6: Add two columns called total and girls_pct.  
Add two columns called total and girls_pct using mutate(), with the total adding together boys and girls and girls_pct being the percentage of the total represented by girls participants. 
```{r}
Participation |> mutate(total = boys+girls) |>
   mutate(girls_pct = (boys-girls)/girls*100)
```

#Task 7: Calculate correlation between boys and girls participation
```{r}
Participation |> summarize(correlation = cor(boys, girls, method="pearson"))
```

#Task 8: Create a scatterplot showing this relationship, adding a line of best fit. Include a headline and label the x and y axes.
```{r}
Participation |> 
  ggplot() +
  geom_point(aes(x=boys, y=girls)) +
  geom_smooth(aes(x=boys, y=girls), method="lm") +
  labs(title="Correlation between the boy and girl participations", x= "Boys", y = "Girls") 

```


# Task 9: In 2-3 sentences, explain what the correlation coefficient and scatterplot reveal about gender equity in Maryland high school sports participation. How do you interpret the school districts that are below the line vs those that are above? Which school districts are most worth examining further, and why? (10 points)

**There is a strong positive correlation between the boy and girl participants in high school sports. This means the more boys participating then the more girls will participate and vise versa. It is important to also consider student body when examining data like this so you are able to compare it to the percentage of girls and boys in the school and see if the percentage is representative of the student body. Overall gender equity in regards to who is participating is pretty good. The districts most worth examining are baltimore county and charles county because those are the ones with the largest gaps in participation among genders.**

## Part 3: Maryland Car Theft Rates (25 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to analyze the data to find out which counties have the highest rates. The fourth dataset contains car theft statistics for Maryland counties in 2023 and population. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv` and complete the following tasks:

#Task 10: Load the car thefts data
```{r}
Car_thefts <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```

#Task 11: Using mutate, add a column that calculates the rate of car thefts for each county - you need to choose the per capita rate 
```{r}
Car_thefts |> mutate(rate_22 = (`2022`/population)*100) |>
   mutate(rate_23 = (`2023`/population)*100)
```

#Task 12: Calculate the median car theft rate and the total number of car thefts statewide. Which counties have rates above the median? 
```{r}
Car_thefts |> summarize(median = median(`2023`), total = sum(`2023`)) 

#The counties with rates above the median are Anne Arundel, Baltimore, Cecil, and Charles.
```


#Task 13: Write 2-3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What's the lede of a story about your findings? 

**The majority of car thefts in Maryland happen in Baltimore County specifically Baltimore City. The lede could be 'Baltimore City has the highest rate of car theft in the state of Maryland.**

## Part 4: Data Analysis Scenario (20 points)

#Task 14: Essay Question.

You receive a tip that local emergency response times have gotten significantly worse over the past year. You obtain monthly data on response times for police, fire and ambulance calls.

Write 3-4 sentences explaining:
1. What statistical measures would you calculate to verify this claim? (10 points)
2. What visualizations would help readers understand the trends? (5 points)
3. What additional context or data would you need to make this a complete story? (5 points)

**I would look up the percentage change of response times over the past three years, then I would compare month by month from the most recent year. This would give me specific data to be able to show what the drop in time has been. I believe a line chart / scatter plot with response times would help readers visualize the trends best. If there is a significant drop they would be able to see the line go down. Additional data to make this a complete story could include employment rates among first responders. If employment has dropped that could explain why response times have gotten slower. **

# Excellent Response


### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your analytical choices
- Submit both your code and written responses
- Load this test to GitHub and submit the link in Elms

Good luck!
