---
title: "Final_Project"
output: html_notebook
---


```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)

```


```{r}
md_crime <- readr::read_csv(
  "https://umd0-my.sharepoint.com/:u:/g/personal/abennet7_umd_edu/EdnEVrxZcYNIlg2r9PDJ_DEB_yaAedY-bKTqtsxwqm3Ngw?download=1"
) |>
  janitor::clean_names()

```

1. The md_county_pop set of data includes the population of all counties in Maryland.
The md_crime set of data includes a list of all crimes that happened in MD throughout 2023-2024. It includes the date & what time the crime occurred, offense name & category, crime against society person or property, and the location. **i have since removed the md_county_pop data as it was breaking my code**
2. The column that will be the most important is county_name. If I am examining crime patterns in counties I will need to be able to seperate what is going on in each county.
3. What isn't necessarily missing, but I want to create new data sets and seperate the crimes by county. Because there is so much data it is a little hard to read currently and that I believe would help me understand what I am looking at more.

4. Producing mean, median, mode, max, min, and standard deviation with incident_hour category below.

```{r}
md_crime$incident_hour <- as.numeric(md_crime$incident_hour)
```

```{r}
mean(md_crime$incident_hour, na.rm = TRUE)
```
```{r}
median(md_crime$incident_hour, na.rm = TRUE)
```

```{r}
max(md_crime$incident_hour, na.rm = TRUE)
```

```{r}
min(md_crime$incident_hour, na.rm = TRUE)
```
```{r}
sd(md_crime$incident_hour, na.rm = TRUE)
#added in part 2 revision
```

Potential hypothesis that I could test include a. the majority of property crimes happen in the evening.
**Revised in pt 2**
Null hypothesis:
The proportion of property crimes that occur when it is dark out, between the hours of 8 p.m. and 6 a.m., is less than or equal to 50%.
H0 : p ≤ 0.50

Alternative hypothesis:
The proportion of property crimes that occur when it is dark out, between the hours of 8 p.m. and 6 a.m., is greater than 50%.
H1 : p > 0.50

```{r}
Property_total <- sum(md_crime$crime_against == "Property", na.rm = TRUE)
```

```{r}
Property_total
```

```{r}
Property_dark <- sum(md_crime$crime_against == "Property" &
    (md_crime$incident_hour >= 20 | md_crime$incident_hour < 6),
    na.rm = TRUE)
```

```{r}
Property_dark
```

```{r}
x <- Property_dark
n <- Property_total
```

```{r}
prop.test(x, n, p = 0.50, alternative = "greater", correct = FALSE)
```

```{r}
binom.test(x, n, p = 0.50, alternative = "greater")
```
**Test Interpretation**
The alternative hypothesis is true because the proportion is greater than 50% and the majority of property crimes do happen when it is dark out. 


```{r}
property_df <- md_crime %>%
  filter(crime_against == "Property")
```

```{r}
property_df$incident_hour <- as.numeric(property_df$incident_hour)
```

```{r}
hour_counts <- property_df %>%
  group_by(incident_hour) %>%
  summarise(count = n(), .groups = "drop")
```

```{r}
ggplot(hour_counts, aes(x = incident_hour, y = count)) +
  geom_col(fill = "steelblue") +
  # Highlight dark hours (20–23 and 0–5)
  annotate("rect", xmin = 20, xmax = 24, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "darkgray") +
  annotate("rect", xmin = 0, xmax = 6, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "darkgray") +
  scale_x_continuous(breaks = 0:23) +
  labs(
    title = "Property Crimes by Hour of Day",
    subtitle = "Shaded region represents dark hours (20:00–06:00)",
    x = "Hour (24-hour format)",
    y = "Number of Property Crimes"
  ) +
  theme_minimal()
```

**Plot note**

I used AI for help generating this plot because I got stuck. Here is the link https://chatgpt.com/s/t_69268702b28c81919f8dff4f1b2a8583. 
