---
title: "Final_Project"
output: html_notebook
---


```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)

```


```{r}
md_crime <- read_csv("~/Desktop/md_crime2023-2024.csv")
```

```{r}
md_county_pop <- read_csv("~/Documents/Jour405A/project/Final_Project/md_county_pop.csv")
```

1. The md_county_pop set of data includes the population of all counties in Maryland.
The md_crime set of data includes a list of all crimes that happened in MD throughout 2023-2024. It includes the date & what time the crime occurred, offense name & category, crime against society person or property, and the location. 
2. The column that will be the most important is county_name. If I am examining crime patterns in counties I will need to be able to seperate what is going on in each county.
3. What isn't necessarily missing, but I want to create new data sets and seperate the crimes by county. Because there is so much data it is a little hard to read currently and that I believe would help me understand what I am looking at more.

4. Producing mean, median, mode, max and min with incident_hour category below.

```{r}
md_crime$incident_hour <- as.numeric(md_crime$incident_hour)
```

```{r}
mean(md_crime$incident_hour, na.rm = TRUE)
```
```{r}
median(md_crime$incident_hour, na.rm = TRUE)
```

```{r}
max(md_crime$incident_hour, na.rm = TRUE)
```

```{r}
min(md_crime$incident_hour, na.rm = TRUE)
```

5. Potential hypothesis that I could test include a. the majority of property crimes happen in the evening or b. Assaults are more common on weekends vs weekdays or c. Shoplifting increased drasticically from 2023 to 2024. or d. Certain types of crimes are more popular in specific seasons 